---
title: "Homework 1 - Glenn Thomakos"
output:
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{r}
# install.packages('dplyr')
library(dplyr)
df_cereal = read.csv('cereal.csv')
summary(df_cereal)
```

First, let's take a quick look at the summary of the cereal dataset. From here, we can have a general idea of what each column represents and whether or not there are any outliers. 

Taking a quick scan of the column titles we can see that some of them will be unhelpful for the goal of identifying a nutrition rating. These unhelpful columns are mfr, type, shelf, and weight.

```{r}
useless_columns = c('mfr','type','shelf','weight')
df_cereal = select(df_cereal, -one_of(useless_columns))
```

From this summary we can also see that one of the columns is "cups" which represents the serving size according to the discussion post on Kaggle for the dataset, found here: https://www.kaggle.com/crawford/80-cereals/discussion/88097. This must be accounted for since people who eat cereal don't measure 0.67 cups of cereal to put in their bowl, they just fill the bowl up. So we change the units for each entry from calories or grams of protein to calories per cup or grams of protein per cup. With this, we also drop the cups column since it's now accounted for in all of the other columns.

```{r}
df_temp = df_cereal[,c('calories','protein','fat','sodium','fiber','carbo','sugars','potass','vitamins')]/df_cereal$cups
df_cereal = cbind(df_cereal[,c('name','rating')],df_temp)
```

Looking back at summary we can also find in the minimums for a few columns the value -1 which is quite odd given you can't have -1g of sugar. A look again at the Kaggle discussion board says that this was simply data not collected. Since this is only 3 cereals I have no problem just eliminating these cereals from the data.

```{r}
missing_value_cereals = c('Quaker Oatmeal','Almond Delight','Cream of Wheat (Quick)')
df_cereal = filter(df_cereal, !(df_cereal$name %in% missing_value_cereals))
```

Now, we can take a look at tables like this that show some egregious outliers. There are some others but this table (along with potass) showed the largest number of outliers. 

```{r}
arrange(df_cereal,desc(fiber))
```

Given the sparse amount, the outliers will also be removed and finally we will have our final data frame for subsequent analysis. 

```{r}
outliers = c('100% Bran', 'All-Bran with Extra Fiber','All-Bran','Grape-Nuts','Great Grains Pecan')
df_cereal = filter(df_cereal,!(df_cereal$name %in% outliers))
```

TODO:
Group together the "good" variables like protein and potassium, then group the "bad" ones as well.

```{r}
set.seed(12)
linear_model = lm(rating~calories+protein+fat+sodium+fiber+carbo+sugars+potass+vitamins, data=df_cereal)
print(summary(linear_model))
```
Here we have the coefficients for each variable that remains as well as some other relevant statistics about the model and we can use this summary to answer question 2.

a: Which predictors appear to have a significant relationship to the response?
We can see that protein, sodium, fiber, and vitamins are the ones with the lowest p-values and are the most significant variables for predicting the health rating while calories is abysmal and offers very little to the model.

b: What does the coefficient variable for "sugar" suggest?
The coefficient suggests that for each unit of sugar in a particular cereal, the model predicts a health rating that is 0.809582 lower than the intercept. For example, if a new cereal was introduced that was literally just 10 units of sugar, the health rating would be about 37.4 (which puts it pretty close to Raisin Bran's 37.8).

```{r}
lm_interactions = lm(rating~calories:sugars+protein+fat+sodium+fiber+sugars+vitamins,data=df_cereal)
summary(lm_interactions)
```
c: Use the * and : symbols to fit models with interactions. Are there any interactions that are significant?
There are. Calories make for a good companion for other features in a "calories per x" scenario which makes sense from a nutritional common sense standpoint. The above is a model that fits the more significant features as well as calories per sugar. Similar models can be made that also feature p-values very near 0 by replacing calories per sugar with calories per sodium and calories per fat.
